---
title: "HiChIP-dataanalysis"
author: "Aditya Mahadevan"
date: "1/7/2022"
output: html_document
---
Here, I will be annotating the anchors and contact regions as hotspots, TSS or enhancers and then will be computing the number of hotspot-hotspot and other interactions. The HiChIP data will be in the form of 2 bed files concatenated side by side and will be doing some pre-processing before using it in R. 
```{r}
#Loading all the required packages
suppressPackageStartupMessages(library(plyranges))
library(GenomicRanges)
library(readxl)
library(tidyverse)
library(dplyr)
library(magrittr)
```


```{r}
# Function to take intersect two bed files using GRanges
intersect_bed <- function(x, y){
  # load into GRanges object
  a <- makeGRangesFromDataFrame(x, keep.extra.columns = T)
  b <- makeGRangesFromDataFrame(y, keep.extra.columns = T)
  # find overlaps
  my_df  <- data.frame(join_overlap_intersect(a,b)) # join_overlap_intersect function from plyranges
}
```


```{r}
#Loading all the required datasets
setwd("~/Documents/HiChIP/datasets")
hotspots = read.delim("PRDM9_BxC_germcells_peaks.txt")
tss = read.delim("mm10_gencode_tss_unique.txt", header = F)
enhancers = read.delim("testis_enhancers.txt", header = F)
hichip = read.delim("5M_k4me3_BxC_allele_interactions.all.mango", header = FALSE)
colnames(hichip) = c("seqnames","start","end","seqnames1","start1","end1")
#hichip$chr = as.character(hichip$chr)
colnames(tss) = c("seqnames","start","end")
colnames(enhancers) = c("seqnames","start","end")
colnames(hotspots) = c("seqnames","start","end")

```


```{r}
## make annotations for anchors
#anchor1 <- hichip %>% mutate(anchor = ifelse(id %in% intersect_bed(hichip[,1:3)], "hotspots", ifelse(tss %in% intersect_bed(,)[,"id"], "tss", "other")))) %>% mutate(anno = factor(anno, levels = c("fdr < 0.01", "within KO region", "QTL target", "Riok3"))) %>% arrange(anno)
## make annotations for contacts

```


```{r}
#subset(hichip[,1:3], (intersect_bed(hichip, tss) == TRUE)) 
#anchor = hichip %>% mutate(anchor = ifelse(subset(hichip[,1:3], (intersect_bed(hichip, hotspots) == TRUE)), "hotspots", ifelse(subset(hichip[,1:3], (intersect_bed(hichip, tss) == TRUE)), "tss")), ifelse(subset(hichip[,1:3], (intersect_bed(hichip, enhancers) == TRUE)), "enhancers", "intergenic"))
```


```{r}
#hichip %>% mutate(style = ifelse(subset(hichip[,1:3], (intersect_bed(hichip, enhancers) == TRUE)),"hot","not")) %>% head()
```


```{r}
#Lets try it another way- I am converting the dataframe to Granges object and then use filter_by_overlaps function of plyranges to look at the overlaps. The idea is to convert the overlap data into dataframe and then add a column that has the information of the dataset which it overlapped it with.
#nrow(subset(hichip[,1:3], (intersect_bed(hichip, tss) == TRUE))) 
#filter_by_overlaps(hichip[,1:3],hotspots)
#test1 = as_granges(hichip[,1:3])
#test2 = as_granges(tss)
#overlap = filter_by_overlaps(test1,test2)
#as.data.frame(overlap) %>% mutate(anchor1="hotspots")

#Creating peak IDs for the hichip dataset
hichip$peakID = paste0("peak_",seq.int(nrow(hichip)))
```
I am getting different numbers of rows with filter_by_overlaps function versus the subset() code. The above code is foolproof so gonna use it real-examples.
I can rbind all the filter_by_overlaps with each dataset and create a dataframe that can be later used for plotting.
```{r}
#Lets convert all the data to Granges
anchors = as_granges(hichip[,c("seqnames", "start", "end", "peakID")])
hotspots_granges = as_granges(hotspots)
tss_granges = as_granges(tss)
enhancers_granges = as_granges(enhancers)
anchors_hotspots = as.data.frame(filter_by_overlaps(anchors,hotspots_granges)) %>% mutate(anchor1="hotspots")
anchors_tss = as.data.frame(filter_by_overlaps(anchors,tss_granges)) %>% mutate(anchor1="tss")
anchors_enhancers = as.data.frame(filter_by_overlaps(anchors,enhancers_granges)) %>% mutate(anchor1="enhancers")
anchors_hichip_overlapping = as_granges(rbind(anchors_hotspots,anchors_tss, anchors_enhancers))
#anchors_hichip_overlapping = unique_granges(as_granges(anchors_hichip_overlapping))
anchors_hichip_nonoverlapping = as.data.frame(filter_by_non_overlaps(anchors, anchors_hichip_overlapping)) %>% mutate(anchor1 = "intergenic")
#anchors_hichip_overlapping = unique(as.data.frame(anchors_hichip_overlapping))
anchors_total = rbind(as_tibble(anchors_hichip_overlapping), anchors_hichip_nonoverlapping)
test1 = anchors_total %>% distinct("seqnames", "start", "end", "width", "strand", "anchor1", .keep_all=TRUE)
```
#
#Just realized that we need anchor1 and anchor2 points side by side; so I have to write a code against the whole hichip dataset.

Similarly generate a dataframe for the contacts
```{r}
hichip[,7:8] = NULL #modify this while modifying the file 
colnames(hichip) = c("seqnames_anchor", "start_anchor", "end_anchor", "seqnames", "start", "end", "peakID")
contacts = as_granges(hichip[,4:7])
contacts_hotspots = as.data.frame(filter_by_overlaps(contacts,hotspots_granges)) %>% mutate(anchor2="HOTSPOTS")
contacts_tss = as.data.frame(filter_by_overlaps(contacts,tss_granges)) %>% mutate(anchor2="TSS")
contacts_enhancers = as.data.frame(filter_by_overlaps(contacts,enhancers_granges)) %>% mutate(anchor2="ENHANCERS")
contacts_hichip_overlapping = as_granges(rbind(contacts_hotspots,contacts_tss, contacts_enhancers))
contacts_hichip_overlapping = (contacts_hichip_overlapping)
contacts_hichip_nonoverlapping = as.data.frame(filter_by_non_overlaps(contacts, contacts_hichip_overlapping)) %>% mutate(anchor2 = "INTERGENIC")
contacts_total = rbind(as.data.frame(contacts_hichip_overlapping), contacts_hichip_nonoverlapping)

#Lets prepare a dataframe that has anchor1(noncapitalized) and anchor2(contact)(which are capitalized) annotation for plotting later
#contacts_hichip_overlapping = as.data.frame(contacts_hichip_overlapping)
#colnames(hichip) = c("seqnames", "start", "end", "seqnames1", "start1", "end1", "peakID")
#hichip_annotation = rbind(contacts_hichip_overlapping, contacts_hichip_nonoverlapping, anchors_hichip_overlapping, anchors_hichip_nonoverlapping)
```
hichip_annotation dataset has all the annotation information in the anchor1 column - with the anchors(non-capitalized) and contacts(capitalized). Some annotations can be overlapping; thats why the hichip_annoation data has little more than twice the number of hichip rows
#hichip annotation dataset has ~twice the number of total peaks as hichip dataset- suggesting the contacts and anchors are there in seperate rows
```{r}
##We need a way to collapse the dataset on the basis of start,end columns and have unique annotation in the anchor1 column
hichip_total = right_join(hichip, anchors_total, by = "peakID") #Not getting the annotations side by side
#colnames(contacts_total) = c("seqnames1", "start1", "end1", "width1", "strand1", "anchor2")
#I have to add the anchor coordinates to join
hichip_total = right_join(hichip_total, contacts_total, by = "peakID")
#names(hichip_total)[7] <- "useless"
```

```{r}
#rough code
#hichip_annotation = hichip_annotation %>% dplyr::filter(!duplicated(cbind(start,end,width,anchor1)))

#Number of anchors and contacts covered
table(hichip_total$anchor1)
table(hichip_total$anchor2)

hichip_total %>% dplyr::group_by(anchor1, anchor2) %>% summarise(counts_contacts = n())

hichip_total %>% dplyr::group_by(anchor1, anchor2) %>% summarise(counts_contacts = n(), groups=c("anchor1", "anchor2")) %>% ggplot(aes(x=anchor1, y=anchor2, size=counts_contacts)) + geom_point() + theme_classic() + xlab("Left anchor") + ylab("Right anchor")  +  labs(size="Number of Overlaps")

#Plotting the numbers only for hotspots
hichip_total %>% dplyr::group_by(anchor1, anchor2) %>% summarise(counts_contacts = n(), groups = c("anchor1", "anchor2")) %>% dplyr::filter(anchor1 == "hotspots") %>% ggplot(aes(x=anchor1, y=anchor2, size=counts_contacts)) + geom_point() + theme_classic() + xlab("Hotspot as left anchor") + ylab("Right anchor")  +  labs(size="Number of Overlaps")
```
Recently, hic data has been performed in individual stages of meiosis. These sort of analyses can be done with those hic datasets and such contacts can be annotated.

1. Proportion, Significance of overlap
2. Bigwig - KaryoplotR
3. ChromHMM
4. Looking at inter-chromosomal interactions
5. CRISPR off SICAP candidates in HEK-P9
6. Intergenic H3K4me3 associated with meiosis datasets and see how they are important in meiosis
HiC pro filtering steps for inter-chromsomal interactions!!!

Bed file of ChromHMM - integrate with the HiChIP data

```{r}
#Read the chromHMM bed file
chromhmm = read.delim("Mouse_11_reordered_dense_chromHMM.bed", header = FALSE)
colnames(chromhmm) = c("seqnames", "start", "end","cluster")
chromhmm = chromhmm[,-c(5:9)] #there are blank rows that might create problems later
```


```{r}
#overlap chromhmm with the anchors_total and contacts_total datasets
#chromhmm_anchor = as.data.frame(filter_by_overlaps(as_granges(chromhmm), as_granges(anchors_total))) %>% mutate(locus = "ANCHORS")
#chromhmm_contacts = as.data.frame(filter_by_overlaps(as_granges(chromhmm), as_granges(contacts_total))) %>% mutate(locus = "CONTACTS")

#chromhmm_total = rbind(chromhmm_anchor, chromhmm_contacts)
#chromhmm_total %>% dplyr::group_by(locus, cluster) %>% summarise(counts_clusters = n())

#Need to get contacts and anchors side by side
#chromhmm$peakClust = paste0("peakClus_",seq.int(nrow(chromhmm)))

#merge the chrom hmm anchors + chrom hmm contacts (chromhmm total) with hichip data
```


```{r}
colnames(hichip_total)
#hichip_total[,1:12] = NULL
chromhmm_co = as.data.frame(pair_overlaps(as_granges(contacts_total), as_granges(chromhmm))) %>% drop_na() %>% dplyr::distinct(peakID, cluster)
chromhmm_an = as.data.frame(pair_overlaps(as_granges(anchors_total), as_granges(chromhmm))) %>% drop_na() %>% dplyr::distinct(peakID, cluster)

chromhmm_total = right_join(chromhmm_co, chromhmm_an, by = "peakID")
```
chromhmm coordinates are large such that multiple annotations are getting mapped to a single location. Therefore, the number of rows in chromhmm total (anchors and contacts information) are larger than individual datasets. 

```{r}
chromhmm_total %>% dplyr::select(cluster.x, cluster.y) %>% dplyr::group_by(cluster.x, cluster.y) %>% dplyr::count() %>% ggplot(aes(x = cluster.x, y = cluster.y, size = n)) + geom_point() + scale_x_discrete(limits = factor(1:11)) + scale_y_discrete(limits = factor(1:11)) + theme_classic() + xlab("Left anchor HMM state") + ylab("Right anchor HMM state")  +  labs(size="Number of Overlaps")

#Plotting the numbers only for hotspots (State 7)
chromhmm_total %>% dplyr::select(cluster.x, cluster.y) %>% dplyr::group_by(cluster.x, cluster.y) %>% dplyr::count() %>% dplyr::filter(cluster.x == 7 | cluster.y == 7) %>% ggplot(aes(x = cluster.x, y = cluster.y, size = n)) + geom_point() + scale_x_discrete(limits = factor(1:11)) + scale_y_discrete(limits = factor(1:11)) + theme_classic() + xlab("Left anchor HMM state") + ylab("Right anchor HMM state") +  labs(size="Number of Overlaps")
                                                              #ggplot(aes(x=anchor1, y=anchor2, size=counts_contacts)) + geom_point() + theme_classic() + xlab("Hotspot as anchor") + ylab("Contact")
#chromhmm_total = chromhmm_total[!duplicated(chromhmm_total$peakID),]
```


```{r}
#colnames(hichip_total)[14:16] = c("seqnames_contact", "start_contact", "end_contact")
#colnames(hichip_total)[1:3] = c("seqnames", "start", "end")
#chromhmm_anchor = as.data.frame(filter_by_overlaps(as_granges(chromhmm), as_granges(hichip_total)))
#rough = left_join(chromhmm_anchor, chromhmm_total, by = c("start", "end"))

#assign peak clust 


#rm(rough)
```



